---
title: "Vital Signs Report"
author: "Robert Jackson"
date: "5/31/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, cache.comments=FALSE,message=FALSE, warning=FALSE}
library("papeR")
library("dplyr")
library("devtools")
library("knitr")
library("ggplot2")
library("kableExtra")
library("corrplot")
library("psych")
library("RMySQL")
```

```{r, echo=FALSE, cache.comments=FALSE,message=FALSE, warning=FALSE}
knitr::opts_chunk$set(root.dir = "/Users/robertjackson/Documents/Classroom_INC./CI_DATA")
```

```{r,echo=FALSE,cache=FALSE, message=FALSE, warning=FALSE}

VS_Skills <- read.csv("SkillDev.VS.5.31.18.csv") # This data for VS_Skills was pulled from MySQL.  The date range is 08.2017 - 04.15.2018.
P_Skills <- read.csv("PointsPos.VS.5.30.18.csv") # This data has no time range.

Grade_Data <- read.csv("user grades.csv") # I have a new dataset which should have more accurate grade levels.
```

```{r,echo=FALSE,cache=FALSE, message=FALSE, warning=FALSE}
Grade_Data <- rename(Grade_Data, Student.ID = user_id)

VS_Skills <- merge(VS_Skills, Grade_Data, by = "Student.ID")
```

```{r,echo=FALSE,cache=FALSE, message=FALSE, warning=FALSE}
# Removeing the ID variable for the Summary Table
VS_Skills <- dplyr::select(VS_Skills, - Student.ID, -Grade)

VS_Skills <- rename(VS_Skills, Grades = gradelevel)
```

***
# Part 1: 
***
```{r,echo = FALSE, cache=FALSE, warning=FALSE}
VS_Skills2 <- na.omit(VS_Skills)
VS_Skills2 <- dplyr::filter(VS_Skills2, episodes > 0)
```

```{r}
VS_Skills2 <- VS_Skills2[!duplicated(VS_Skills2), ]
```

### Summary Statistics of Vital Signs - 2017 - 18 [table 1]
```{r,echo = FALSE, cache=FALSE, warning=FALSE}
# Defining the function
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}

# identifying numeric columns
ind <- sapply(VS_Skills2, is.numeric)


# applying the function to numeric columns only
sapply(VS_Skills2[, ind], my.summary)  
```

### Histograms of Skill Point Distribution *[Cummunlative]*

```{r,histogram, echo=FALSE, cache=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) + geom_histogram(mapping = aes(x = Social), binwidth = 0.5) + ggtitle("Skills Performance in Vital Signs", subtitle = "Social & Cross-Cultural") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 16,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14
  )) + geom_vline(xintercept =
  mean(VS_Skills2$Social) , color = "green")
```

```{r,echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = Productivity), binwidth = 0.5) + ggtitle("Skills Performance in Vital Signs", subtitle = "Productivity & Accountability") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 16,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14
  )) + geom_vline(xintercept =
  mean(VS_Skills2$Productivity) , color = "green")
```

```{r,echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = Leadership), binwidth = 0.5) + ggtitle("Skills Performance in Vital Signs", subtitle = "Leadership & Responsibility") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 16,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14
  )) + geom_vline(xintercept =
  mean(VS_Skills2$Leadership) , color = "green")
```


```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = Initiative), binwidth = 0.5) + ggtitle("Skills Performance in Vital Signs", subtitle = "Initiative & Self-Direction") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 16,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14
  )) + geom_vline(xintercept =
  mean(VS_Skills2$Initiative) , color = "green")
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = Flexibility), binwidth = 0.5) + ggtitle("Skills Performance in Vital Signs", subtitle = "Flexibility & Adaptability") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 16,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14
  )) + geom_vline(xintercept =
  mean(VS_Skills2$Flexibility) , color = "green")
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = Total.Good.Decisions), binwidth = 0.5) + ggtitle("Performance in Vital Signs", subtitle = "Total Good Decision") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 16,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14
  )) + geom_vline(xintercept =
  mean(VS_Skills2$Total.Good.Decisions) , color = "green")
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = episodes), binwidth = 0.5) + ggtitle("Distribution of Episodes Played in Vital Signs") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = ,
  size = 12
  )) + geom_vline(xintercept =
  mean(VS_Skills2$episodes) , color = "green")
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
ggplot(data = VS_Skills2) +
  geom_histogram(mapping = aes(x = Grades), binwidth = 0.5) + scale_x_continuous(,breaks = 1:13, limits=c(0, 14)) + ggtitle("Distribution of Student Grade Levels who Play Vital Signs",subtitle = "GREEN LINE IS MEDIAN") + labs(x = "Total Score", y = "Number of Students") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14,
  hjust = 0
  )) +
  theme(axis.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = ,
  size = 12
  )) + geom_vline(xintercept =
  median(VS_Skills2$episodes) , color = "green")
```
### Summary Statistics of Skill Point Distribution [By Percentage]
*The additional varaibles takes into account the particular episodes played by students and which points were available for them to score.*

```{r,echo=FALSE,cache=FALSE, message=FALSE, warning=FALSE}
VS_Skills2 <- dplyr::rename(VS_Skills2, Users = User_Name)
P_Skills <- dplyr::rename(P_Skills, Users = username)
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
# Filter to remove students who have completed 0 episodes
VSPer_Skills <- merge(P_Skills, VS_Skills2, by = "Users")
VSPer_Skills <- dplyr::filter(VSPer_Skills, episodes > 0)
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
VSPer_Skills <-
  dplyr::select(VSPer_Skills,
  -Social,
  -Initiative,
  -Leadership,
  -Productivity,
  -Flexibility)
```

```{r, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
# Need to generate the ratio necessary accurately understand the data.  I have to make new variables based on the ratio

VSPer_Skills <-
dplyr::mutate(VSPer_Skills, SocialR = student_good_social / possible_good_social)
VSPer_Skills <-
dplyr::mutate(VSPer_Skills, ProdR = student_good_Productivity / possible_good_Productivity)
VSPer_Skills <-
dplyr::mutate(VSPer_Skills, LeadR = student_good_Leadership / possible_good_Leadership)
VSPer_Skills <-
dplyr::mutate(VSPer_Skills, InitiativeR = student_good_Initiative / possible_good_Initiative)
VSPer_Skills <-
dplyr::mutate(VSPer_Skills, FlexR = student_good_Flexibility / possible_good_Flexibility)
```

```{r, echo=FALSE, cache.comments=FALSE,message=FALSE, warning=FALSE,results = 'hide'}

# Some how all of the NAs that exisited in past data iterations are gone.  I'm going to need to investigate.

#          Users        student_good_social       possible_good_social  student_good_Productivity 
#                        0                          0                          0                          0 
#possible_good_Productivity    student_good_Leadership   possible_good_Leadership    student_good_Initiative 
#                         0                          0                          0                          0 
#  possible_good_Initiative   student_good_Flexibility  possible_good_Flexibility       Total.Good.Decisions 
#                         0                          0                          0                          0 
#                  episodes                     Grades                    SocialR                      ProdR 
#                         0                          0                          1                          0 
#                     LeadR                InitiativeR                      FlexR 
#                         0                          1                          0 
                         
sapply(VSPer_Skills, function(x)
sum(is.na(x)))
```

```{r,echo = FALSE, cache=FALSE, warning=FALSE}
VSPer_Skills2 <- select (VSPer_Skills, SocialR,ProdR,LeadR,InitiativeR,FlexR)
```

### Motified Summary Statistics of Vital Signs Including Percentage - 2017 - 18 [table 2]
```{r,echo = FALSE, cache=FALSE, warning=FALSE}
# Defining the function
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}

# identifying numeric columns
ind <- sapply(VSPer_Skills2, is.numeric)


# applying the function to numeric columns only
sapply(VSPer_Skills2[, ind], my.summary)
```

```{r,echo = FALSE, cache=FALSE, warning=FALSE}
VSPer_Skills3 <- VSPer_Skills
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Earned_Social = student_good_social)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Earned_Productivity = student_good_Productivity)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Earned_Leadership = student_good_Leadership)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Earned_Flexibility = student_good_Flexibility)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Earned_Initiative = student_good_Initiative)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Available_Social = possible_good_social)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Available_Productivity = possible_good_Productivity)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Available_Leadership = possible_good_Leadership)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Available_Flexibility = possible_good_Flexibility)
VSPer_Skills3 <-
dplyr::rename(VSPer_Skills3, Points_Available_Initiative = possible_good_Initiative)
```

```{r,echo = FALSE, cache=FALSE, warning=FALSE}
VSPer_Skills4 <-
  select(
  VSPer_Skills3,
  Points_Earned_Social,
  Points_Earned_Productivity,
  Points_Earned_Leadership,
  Points_Earned_Flexibility,
  Points_Earned_Initiative,
  Points_Available_Social,
  Points_Available_Productivity,
  Points_Available_Leadership,
  Points_Available_Flexibility,
  Points_Available_Initiative
  )
```

```{r,echo = FALSE, cache=FALSE, warning=FALSE}
# Defining the function
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}

# identifying numeric columns
ind <- sapply(VSPer_Skills4, is.numeric)


# applying the function to numeric columns only
sapply(VSPer_Skills4[, ind], my.summary)
```

### Histogram of Skill Point Distribution [By Percentage]
*This takes into account the particular episodes played by students and which points were avaiable for them to score.  The distribution is now a scale of 0 - 100%.*  

*Meaning student could have earns 0% of the points available to them based on the episodes they played or they may have earned 100% of the point available to them in the episodes they've played.*

### Box Plot of 21st Century Skills by Percentage

```{r,echo=FALSE}
VSPer_Skills2 <- rename(VSPer_Skills2, "Flexiblity" = "FlexR") #Flexibility & Adaptability
VSPer_Skills2 <- rename(VSPer_Skills2, "Initiative" = "InitiativeR") #Initiative & Self-Direction 
VSPer_Skills2 <- rename(VSPer_Skills2, "Productivity" = "ProdR") #Productivity & Accountability
VSPer_Skills2 <- rename(VSPer_Skills2, "Leadership" = "LeadR") #Leadership & Responsibility
VSPer_Skills2 <- rename(VSPer_Skills2, "Social" = "SocialR") #Social & Cross-Cultural Skills 
```

```{r,echo = FALSE, cache=FALSE, warning=FALSE}
library("reshape2")
```

```{r,echo=FALSE}
Viz <- melt(VSPer_Skills2)
```


```{r, echo=FALSE}
ggplot(data = Viz, aes(x = variable, y = value, fill = variable))  + geom_boxplot() + scale_fill_discrete(
  name = "21st Century Skills",
  labels = c(
  "Social & Cross-Cultural",
  "Productivity & Accountability",
  "Leadership & Responsibility",
  "Initiative & Self-Direction",
  "Flexibility & Adaptability"
  )
  ) + theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank()
  ) + stat_summary(
  fun.y = mean,
  geom = "point",
  shape = 20,
  size = 5,
  color = "green",
  fill = "green"
  ) + stat_summary(
  fun.y = median,
  fun.ymin = median,
  fun.ymax = median,
  geom = "crossbar",
  width = 0.9
  
  ) + ggtitle("Points Earned by Percentage - Vital Signs", subtitle = "Measures of Five 21st Century Skills") + labs(x = "Skills", y = "Percentage Earned") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14,
  hjust = 0
  ))
```

**We'll use this next section to outline how many student had an opportunity to perform each of the five 21st century skills**

```{r}
D1 <- dplyr::select(VSPer_Skills, contains("possible_"))
```

```{r,echo=FALSE}
Viz2 <- melt(D1)
```

**This boxplot is dynamic and analyzes what episodes of the game students reach and whether they encounter an opportunity to demonstrate one of the five 21st century skills.  Since these skills don't appear evenly throughout the game its important to know in which ways students are not being exposed to certain skill opportunities.**
```{r, echo=FALSE}
ggplot(data = Viz2, aes(x = variable, y = value, fill = variable)) + scale_fill_discrete(
  name = "21st Century Skills Available",
  labels = c(
  "Social & Cross-Cultural",
  "Productivity & Accountability",
  "Leadership & Responsibility",
  "Initiative & Self-Direction",
  "Flexibility & Adaptability"
  )
  ) + geom_boxplot() + theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank()
  ) + stat_summary(
  fun.y = mean,
  geom = "point",
  shape = 20,
  size = 5,
  color = "green",
  fill = "green"
  ) + stat_summary(
  fun.y = median,
  fun.ymin = median,
  fun.ymax = median,
  geom = "Crossbar",
  width = 0.9
  ) + ggtitle("Range of Points Available - Vital Signs", subtitle = "Measures of Five 21st Century Skills") + labs(x = "Skills", y = "Points Available to Earn") + theme(plot.title = element_text(
  family = "Arial",
  color = "#a4c94f",
  face = "bold",
  size = 14,
  hjust = 0
  ))
```

```{r,echo = FALSE, cache=FALSE, warning=FALSE}
P1 <- dplyr::select(VSPer_Skills, contains("possible_"))
P1 <- 
  dplyr::rename(P1, Points_Available_Social = possible_good_social)
P1 <-
  dplyr::rename(P1, Points_Available_Productivity = possible_good_Productivity)
P1 <-
  dplyr::rename(P1, Points_Available_Leadership = possible_good_Leadership)
P1 <-
  dplyr::rename(P1, Points_Available_Flexibility = possible_good_Flexibility)
P1 <-
  dplyr::rename(P1, Points_Available_Initiative = possible_good_Initiative)
```

### Table of Points Availabe to Earn - Vital Signs
**These are the count of students who have encountered an opportunity to demonstrate a particular 21st century skill. **
```{r,echo = FALSE, cache=FALSE, warning=FALSE}
colSums(P1 != 0)
```


***
# Part Two: 

***

*We're now going to add the survey data.  Go through the same EDA. Then finally we'll look at the correlation between the skills in VS and the measures in the survey data.*

```{r, echo=FALSE}
# This data is the majority of schools and students
Survey1 <-
read.csv("SY 2017-2018 Student Pre-Program Survey_January 29, 2018_11.38.csv")
# This data is from one school/ system in Milwaukee
Survey2 <-
read.csv("SY 2017-2018 MPS Student Pre-Program Survey_January 29, 2018_11.34.csv")
```

```{r, echo=FALSE}
Survey1a <- dplyr::select(Survey1, Q21_1:Q21_10)
Survey2a <- dplyr::select(Survey2, Q22_1:Q22_10)
```

### Working through Survey 1

```{r, echo=FALSE, results= 'hide'}
summary(Survey1a)
```

```{r, echo=FALSE, results = 'hide'}
# no NA values

sapply(Survey1a, function(x)
sum(is.na(x)))
```

```{r, echo=FALSE}
# I need to remove the beginning part of all statements in row 2.  Then I need to change the variable name to fit the question.

Survey1a <- data.frame(lapply(Survey1a, function(x) {
gsub("Please respond to the following statements. - ", "", x)
}))

Survey2a <- data.frame(lapply(Survey2a, function(x) {
gsub("Please respond to the following statements. - ", "", x)
}))
```

```{r, echo=FALSE}
Survey1a = Survey1a[-1,]
Survey1a = Survey1a[-1,]          
```

```{r, echo=FALSE}
#I've been running into an issue converting the rows form factors into integers.  This is important for properly running any correlation.  I suspect that the high numbe of "blank" rows plays a role here.  I'll try to convert those into a character/or then remove them.

#We'll have to keep track of how many rows we'll lose before we can jsut remove data.
# Converting the results of the external survey into integars from factors so that I can use them in the correlation matrix.
Survey1b <-
data.frame(lapply(Survey1a, function(x)
as.numeric(as.character(x))))
```

```{r, echo=FALSE, results= 'hide'}
# All the blanks were converted into 0s so i haven't lost any rows of data.
which(Survey1b$Q21_1 == 67)
which(Survey1b$Q21_1 == 88)
```

```{r, echo=FALSE, results= 'hide'}
row_sub = apply(Survey1b, 1, function(row)
  all(row != 0))
  ##Subset as usual
  sapply(Survey1b, function(x)
  sum(is.na(x)))
```

```{r,echo=FALSE, results= 'hide'}
# removing the na's
Survey1b <- na.omit(Survey1b)
```

```{r, echo=FALSE}
# Note in this convertion processs we lost a little under 400 rows of data.  This is due to the Na's
A1 <- cor(Survey1b)
```

**How the Survey questions correlate**
```{r,echo=FALSE}
corrplot(
  A1,
  type = "lower",
  order = "hclust",
  tl.col = "black",
  method = "number"
  ) #plot matrix
  
  corrplot(
  A1,
  type = "lower",
  order = "hclust",
  tl.col = "black",
  method = "circle"
  ) #plot matrix
```

**Survey 2**

```{r, echo=FALSE, results= 'hide'}
sapply(Survey2a, function(x)
  sum(is.na(x)))
```

```{r, echo=FALSE,results= 'hide'}
Survey2a = Survey2a[-1,]
Survey2a = Survey2a[-1,]         
```

```{r, echo=FALSE}
# Converting the results of the external survey into integars from factors so that I can use them in the correlation matrix.
Survey2b <-
  data.frame(lapply(Survey2a, function(x)
    as.numeric(as.character(x))))
```

```{r,echo=FALSE, results= 'hide'}
# All the blanks were converted into 0s so i haven't lost any rows of data.
which(Survey2b$Q21_1 == 67)
which(Survey2b$Q21_1 == 88)
```

```{r, echo=FALSE, results= 'hide'}
row_sub = apply(Survey2b, 1, function(row)
  all(row != 0))
  ##Subset as usual
  sapply(Survey2b, function(x)
  sum(is.na(x)))
```

```{r, echo=FALSE}
# removing the na's
Survey2b <- na.omit(Survey2b)
```

```{r, echo=FALSE}
# Note in this convertion processs we lost a little under 400 rows of data.  This is due to the Na's
B1 <- cor(Survey2b)
```

### Corplot of the Milwaukee Survey Questions

```{r, echo=FALSE}
corrplot(
  B1,
  type = "lower",
  order = "hclust",
  tl.col = "black",
  method = "number"
  ) #plot matrix
```

```{r, echo=FALSE}
  corrplot(
  B1,
  type = "lower",
  order = "hclust",
  tl.col = "black",
  method = "circle"
  ) #plot matrix
```

***
# Part Three: Creating the 2 cumulative measures from the External Instruments & Time to merge External and internal metrics

***
```{r, echo=FALSE}
# This data is the majority of schools and students
Survey1 <-
read.csv("SY 2017-2018 Student Pre-Program Survey_January 29, 2018_11.38.csv")
# This data is from one school/ system in Milwaukee
Survey2 <-
read.csv("SY 2017-2018 MPS Student Pre-Program Survey_January 29, 2018_11.34.csv")
```


```{r, echo=FALSE, results='hide'}
match(3,Survey1$Q11)
match("3,",Survey1$Q11)
```

```{r, echo=FALSE}
# Cutting out the variables are not useful for our purpose regarding correlation
Survey1a <- dplyr::select(Survey1, ResponseId:Q21_10)
Survey2a <- dplyr::select(Survey2, ResponseId:Q22_10)
```

```{r, echo=FALSE}
Survey1a <- dplyr::select(Survey1a, ResponseId, Q8 , Q21_1:Q21_10)
Survey2a <- dplyr::select(Survey2a, ResponseId, Q11, Q22_1:Q22_10)
```

```{r, echo=FALSE}
Survey1a <- data.frame(lapply(Survey1a, function(x) {
  gsub("Please respond to the following statements. - ", "", x)
}))

Survey2a <- data.frame(lapply(Survey2a, function(x) {
gsub("Please respond to the following statements. - ", "", x)
}))
```

```{r, echo=FALSE}
colnames(Survey1a) <- as.character(unlist(Survey1a[1, ]))
Survey1a = Survey1a[-1,]

colnames(Survey2a) <- as.character(unlist(Survey2a[1, ]))
Survey2a = Survey2a[-1,] 
```

```{r, echo=FALSE}
Survey1a = Survey1a[-1,]
Survey2a = Survey2a[-1,] 
```

```{r,echo=FALSE}
#Merging the two data frames
SurveyM <- rbind(Survey1a, Survey2a)
```

```{r}
# Renaming the Variables for ease of use.

SurveyM2 <- dplyr::rename(
SurveyM,
"Users" = "Your Game Username",
"Q1" = "I can set an example for others.",
"Q2" = "I can contribute to group efforts.",
"Q3" = "I can take an active role in learning.",
"Q4" = "I compare possible solutions to find the best one.",
"Q5" = "I can do things on my own.",
"Q6" = "I usually have more than one source of information before making a decision.",
"Q7" = "I take steps to accomplish goals.",
"Q8" = "I can be in charge of a group of classmates.",
"Q9" = "I compare different ideas when thinking about a topic.",
"Q10" = "When faced with a decision, I realize that some choices are better than others.",
"Q11" = "I use what I know already to solve new problems.",
"Q12" = "I think about all the information I have about making new decisions."
)
```

```{r, echo=FALSE}
# Converting the survey responses to numeric so we can have a cummulative score
SurveyM2$Q1 <- as.numeric(as.character(SurveyM2$Q1))
SurveyM2$Q2 <- as.numeric(as.character(SurveyM2$Q2))
SurveyM2$Q3 <- as.numeric(as.character(SurveyM2$Q3))
SurveyM2$Q4 <- as.numeric(as.character(SurveyM2$Q4))
SurveyM2$Q5 <- as.numeric(as.character(SurveyM2$Q5))
SurveyM2$Q6 <- as.numeric(as.character(SurveyM2$Q6))
SurveyM2$Q7 <- as.numeric(as.character(SurveyM2$Q7))
SurveyM2$Q8 <- as.numeric(as.character(SurveyM2$Q8))
SurveyM2$Q9 <- as.numeric(as.character(SurveyM2$Q9))
SurveyM2$Q10 <- as.numeric(as.character(SurveyM2$Q10))
SurveyM2$Q11 <- as.numeric(as.character(SurveyM2$Q11))
SurveyM2$Q12 <- as.numeric(as.character(SurveyM2$Q12))
```

```{r, echo=FALSE}
# We're creating two variables. They effectively take into account the external measures.
SurveyM2$LCSDI <-
  (SurveyM2$Q1 + SurveyM2$Q2 + SurveyM2$Q3 + SurveyM2$Q5 + SurveyM2$Q7 + SurveyM2$Q8/6)

SurveyM2$CTDMPS <-
  (SurveyM2$Q4 + SurveyM2$Q6 + SurveyM2$Q9 + SurveyM2$Q10 + SurveyM2$Q11 + SurveyM2$Q12/6)
```

```{r,echo=FALSE}
# I need to merge the questions that match particular skill. They need to be averages since this is on a Likert scale.  

SurveyM2$L_C <-
  (SurveyM2$Q1 + SurveyM2$Q2 + SurveyM2$Q8/3) # L/C = Leadership / Collaboration
SurveyM2$SD_I <-
  (SurveyM2$Q3 + SurveyM2$Q7 + SurveyM2$Q5/3) # SD/I = Self Direction/ Initiative
SurveyM2$CT <- (SurveyM2$Q6 + SurveyM2$Q9/2) # CT = Critical Thinking
SurveyM2$DM <- (SurveyM2$Q12 + SurveyM2$Q10/2) # DM = Decision Making
SurveyM2$PS <- (SurveyM2$Q11 + SurveyM2$Q4/2) # PS = Problem Solving
```

```{r, echo=FALSE}
# Removing the questions
SurveyM2 <-
  dplyr::select(SurveyM2,
  -Q1,
  -Q2,
  -Q3,
  -Q4,
  -Q5,
  -Q6,
  -Q7,
  -Q8,
  -Q9,
  -Q10,
  -Q11,
  -Q12)
```

```{r, echo=FALSE}
# clearning up the username issues/ inconsistencies 
A1 <-
  VSPer_Skills3 #Vital Signs Data from 08.15.2017 - Present
B1 <-
  read.csv("SY17-18.All.TestComplete.csv") #Data that has accurate User Names
```

```{r, echo=FALSE}
# Eliminating the variables that I don't need
B1 <- dplyr::select(B1, Student, STAR.RTL.ID, Student.Survey.Match)
```

```{r, echo=FALSE}
#Need to merge the survey response data from SurveyM2 with the accurate User Name & ID data from B1
B2 <- dplyr::rename(B1, Response_ID = Student.Survey.Match)
SurveyM3 <- dplyr::rename(SurveyM2, Response_ID = "Response ID")
B3 <- merge(B2, SurveyM3, by = "Response_ID")
```

### Summary Statistics of External Measures - 2017 - 18 [table 3]
```{r, echo = FALSE,cache=FALSE,warning=FALSE, results = 'asis'}
# merge of the the Survey Data w/ the two measures and their accurate Usernames
B3_Summary <- dplyr:: select(B3, - Users, - STAR.RTL.ID, -Student, -Response_ID)

# Defining the function
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}

# identifying numeric columns
ind <- sapply(B3_Summary, is.numeric)


# applying the function to numeric columns only
sapply(B3_Summary[, ind], my.summary) 
```

```{r,echo=FALSE}
B4 <- dplyr::select(B3, - Users)
B4 <- dplyr::rename(B4, Users = STAR.RTL.ID)
MasterFrame1 <- merge(VSPer_Skills3, B4, by="Users" )
```

```{r}
write.csv(MasterFrame1, file = "AllVariables.VS.6.5.2018.csv")
```

MasterFrame1 <- dplyr::select(MasterFrame1, Users, Total.Good.Decisions, episodes, SocialR, ProdR, LeadR, InitiativeR, FlexR, Response_ID, Student, LCSDI, CTDMPS, L_C, SD_I, CT, DM, PS)

### Summary Statistics of Vital Signs & Survey Data 2017 - 2018 [table 4]
```{r,echo = FALSE,cache=FALSE,warning=FALSE, results = 'asis'}
# Defining the function
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}

# identifying numeric columns
ind <- sapply(MasterFrame1, is.numeric)


# applying the function to numeric columns only
sapply(MasterFrame1[, ind], my.summary) 
```

```{r, echo = FALSE,cache=FALSE,warning=FALSE, results = 'asis'}
Master_Summary <- dplyr::select(MasterFrame1, Total.Good.Decisions, episodes, SocialR, ProdR, LeadR, InitiativeR, FlexR, Response_ID, Student, LCSDI, CTDMPS, L_C, SD_I, CT, DM, PS)
```

```{r, echo=FALSE}
# Converting the results of the external survey into integars from factors so that I can use them in the correlation matrix.
M1 <- dplyr::select(Master_Summary,-Response_ID,-Student)

M2 <- na.omit(M1)

MCor <- cor(M2)
```

## Correlation plot of the External measures and the Internal measures from VS
```{r, echo=FALSE}
corrplot(MCor,type = "lower", order = "hclust", tl.col = "black", method = "number") #plot matrix
```

```{r, echo=FALSE}
corrplot(MCor,type = "lower", order = "hclust", tl.col = "black", method = "circle") #plot matrix
```

***
# Part 4: PCA/ Factor Analysis
***

```{r}
M3 <- dplyr::select(M2, -Total.Good.Decisions, - episodes)
```


```{r, echo=FALSE}
#principal component analysis
prin_comp <- prcomp(M3, scale. = T)
names(prin_comp)
```

```{r, echo=FALSE}
prin_comp$rotation
biplot(prin_comp, scale = 0)
summary(prin_comp)
```

```{r, echo=FALSE}
std_dev <- prin_comp$sdev

#compute variance
pr_var <- std_dev ^ 2
```

```{r,echo=FALSE}
#proportion of variance explained
prop_varex <- pr_var / sum(pr_var)
prop_varex[1:20]
```

```{r, echo=FALSE}
#scree plot
plot(prop_varex,
     xlab = "Principal Component",
     ylab = "Proportion of Variance Explained",
     type = "b")
```

```{r,echo=FALSE}
#cumulative scree plot
plot(cumsum(prop_varex),
     xlab = "Principal Component",
     ylab = "Cumulative Proportion of Variance Explained",
     type = "b")
```

**Rerunning the PCA with a different package**
```{r,echo=FALSE}
library("ggplot2")
library("FactoMineR")
library("factoextra")
```

```{r,echo=FALSE}
# The R code below, computes principal component analysis on the active individuals/variables:

# the eigenvalues measure the amount of variation retained by each principal component. Eigenvalues are large for the first PCs and small for the subsequent PCs. That is, the first PCs corresponds to the directions with the maximum amount of variation in the data set.

# The sum of all the eigenvalues give a total variance of 10.
M2.pca <- PCA(M3, graph = FALSE)
print(M2.pca)
```

```{r,echo=FALSE}
eig.val <- get_eigenvalue(M2.pca)
eig.val
```

```{r,echo=FALSE}
fviz_eig(M2.pca, addlabels = TRUE, ylim = c(0, 50))
```

```{r, echo=FALSE}
var <- get_pca_var(M2.pca)
var
```

```{r, echo=FALSE}
# Coordinates of variables
head(var$coord, 9)
```

```{r, echo=FALSE}
fviz_pca_var(M2.pca, col.var = "black")
```

```{r, echo=FALSE}
corrplot(var$cos2, is.corr=FALSE)
```

## How the variables cluster together
```{r, echo=FALSE}
# Create a grouping variable using kmeans
# Create 2 groups of variables (centers = 2)
set.seed(123)
res.km <- kmeans(var$coord, centers = 2, nstart = 25)
grp <- as.factor(res.km$cluster)
# Color variables by groups
fviz_pca_var(
M2.pca,
col.var = grp,
palette = c("#0073C2FF", "#EFC000FF", "#868686FF"),
legend.title = "Cluster"
)
```

```{r, echo=FALSE}
M2.desc <- dimdesc(M2.pca, axes = c(1, 2), proba = 0.05)
# Description of dimension 1
M2.desc$Dim.1
```

```{r, echo=FALSE}
# Description of dimension 2
M2.desc$Dim.2
```

***
# Part 5: Regression

***

**Regression LCSDI**

_the varibles that make up majority of PC1 as it relates to LCSDI (External1)_
I will run a regression on the variables taht have over 75% correlation to PC1

```{r, echo=FALSE}
External1.a <-
  lm(LCSDI ~  +FlexR + Total.Good.Decisions + episodes, data = M2)
  summary(External1.a)
```

```{r, echo=FALSE}
External1.b <-
  lm(LCSDI ~  SocialR + FlexR + Total.Good.Decisions + episodes, data = M2)
  summary(External1.b)
```

_the varibles that make up majority of PC1 as it relates to CTDMPS (External2)_
I will run a regression on the variables taht have over 75% correlation to PC1
```{r, echo=FALSE}
External2.a <-
  lm(CTDMPS ~  +FlexR + Total.Good.Decisions + episodes, data = M2)
  summary(External2.a)
```

```{r, echo=FALSE}
External2.b <-
  lm(CTDMPS ~ SocialR + FlexR + Total.Good.Decisions + episodes, data = M2)
  summary(External2.b)
```










